Towards completely caring 15-minute neighbourhoods

### 2024-10-10 paper/Manuscript.tex
Submitted to journal with editorial correspondance. With lock file update.

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{unicode-math}
\usepackage{hyperref}
\def\tightlist{}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Caring 15-minute neighbourhoods},
  pdfauthor={Anastasia Soukhov; Léa Ravensbergen; Lucía Mejía Dorantes; Antonio Páez},
  pdfkeywords={Accessibility, Mobility of Care, Gender, 15-minute
neighbourhood, Care Complete Neighbourhoods},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Caring 15-minute neighbourhoods}
\author[1]{Anastasia Soukhov%
\corref{cor1}%
}
 \ead{soukhoa@mcmaster.ca} 
\author[1]{Léa Ravensbergen%
%
}
 \ead{ravensbl@mcmaster.ca} 
\author[2]{Lucía Mejía Dorantes%
%
}
 \ead{lucia.mejia.dorantes@gmail.com} 
\author[1]{Antonio Páez%
%
}
 \ead{paezha@mcmaster.ca}

@@

The ``15-Minute City'' concept has been embraced by global leaders to
promote human-scale neighbourhoods with transport and land-use designs
that support short trips to daily necessities. This paper bridges the
15-Minute City to ``Mobility of Care'', a framework that foregrounds
travel to care destinations, travel done predominately by women. This
focus contrasts the more commonly studied travel to employment and
leisure destinations. While the 15-Minute City concept is flexible
enough to consider all destination types, gendered examinations are
relatively lacking in the literature, and little research to date has
focused explicitly on care destinations. This gap is addressed in this
paper by identifying which areas in the city of Hamilton, Canada are
`caring 15-minute neighbourhoods'. To do so, a database of care
destinations was compiled to estimate the number (using the cumulative
opportunities accessibility measure) and diversity of mix (using the
entropy measure) of care destinations within a 15-minute walk from
residential parcels. Using data-driven machine learning techniques
(Self-Organizing Maps and Decision Trees), neighborhoods were classified
into `caring 15-minute neighbourhood' typologies that are examined
across residential socio-economic profiles. Our results suggest that the
majority of caring 15-minute neighbourhoods are in the urban core, where
more lower-income households reside. In contrast, not caring 15-minute
neighbourhoods in higher-income peripheral areas. Areas that make good
candidates for urban policy intervention are identified, and the
implications of enhancing 15-minute walkable caring access are discussed
in relation to gender mainstreaming in transportation planning.

Keywords: Accessibility; 15-Minute Neighborhoods; Chrono-Urbanism;
Cities of Proximity; Mobility of Care; Inequality; Gender Gap;

@@

The term 15-Minute City is relatively new but the concept is not, as it
is simply a normative cumulative-opportunities accessibility measure
\citep{paez_measuring_2012}. Its newfound fame under the moniker of
chronourbanism comes to complement long-standing efforts to foster
community and local travel to amenities in neighbourhood planning
practice. As recent examples, planned neighbourhoods dominated post-WWII
built urban form: the Neighbourhood Unit Concept in the western world
\citep{brodyNeighbourhoodUnitConcept2013} and a parallel concept of the
``mikrorayons'' neighbourhoods in the Soviet Union
\citep{kissfazekasCircleParadigms15minute2022}. Along different
dimensions, these planning forms have been extensively critiqued
\citep{talenSocialSciencePlanned2017}. In this respect, neighbourhood
planning approaches such as the 15-Minute City concept offer an
opportunity for bottom-up approaches that leverage co-creation tools and
meaningful resident participation to achieve equitable and just
neighbourhood forms \citep{mahmoudCocreationPathwayUrban2021}. How to
prescribe equitable urban forms that `enable' contact with social
opportunities (instead of `engendering it' from the top-down) is still a
matter for debate. Questions remain, including: what destinations really
matter? What mode and travel time threshold is most appropriate? And who
benefits? To be certain, these dimensions are under discussion among
proximity-based planners who use accessibility-based tools
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022, guzmanProximityEnoughCritical2024}.

From another perspective, feminist geographers, urban planners and other
researchers agree that urban structure influences travel, but city
planning falls short in planning for a neutral identity. In contrast,
cities ought to be planned according to the multiple identities of the
inhabitants
\citep{vacchelli_towards_2018, urbandevelopmentviennaGenderMainstreamingUrban2013}.
Researchers have demonstrated how mobility patterns differ according to
gender
\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990, neto_gender_2015}.
The causes for this are varied, and related to gendered social norms and
cultural factors that play a role in how, where, with whom, what for,
and how far we move. Of note is that women tend to travel more
frequently and for shorter distance
\citep{roorda_triap_2010, morency_distance_2011} and have more
activities related to care \citep{ilo_care_2018} no matter at what life
stage they are \citep{garciaromanGenderDifferencesTime2022}. Combined,
these findings indicate that the spatial organization of activities has
a different impact according to gender, and therefore the 15-Minute City
concept should not be gender blind.

This work builds a theoretical bridge between the 15-Minute City and
mobility of care concepts to answer questions about what destinations
matter and who benefits from proximity planning. The Mobility of Care
term was coined in
\citet{sanchezdemadariagaMobilityCareIntroducing2013}; it refers to all
travel required to sustain the needs of a household, such as grocery
shopping, escorting children, travelling to health appointments, and
running errands. While decades of research have examined types of
household-serving trips (e.g., shopping trips, escorting to school),
\citet{sanchezdemadariagaMobilityCareIntroducing2013} was the first to
consider them all as one category. In doing so, care-related mobility
can be shown to comprise a significant proportion of daily travel,
characterizing approximately 30\% of adults' daily trips
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
We argue that Mobility of Care is especially relevant to the 15-Minute
City, as care trips are necessary for \emph{all} people and they most
often occur at the local level. Moreover, care trips are of particular
gendered and socio-economic import, as women, and especially those in
lower income households, complete the majority of care trips
\citep{ravensbergen2023exploratory}. Despite care travel's importance in
advancing gender equality -and equity more broadly-, mobility of care
and access to care destinations have been under-examined relative to
employment destinations. This is especially pertinent in accessibility
research that has largely focused on employment points of interest e.g.,
\citep{farberOntarioLineSocioeconomic2019, duarteInfluenceJobAccessibility2023, ryanAccessibilitySpaceTime2023}.

@@

In the last decades, the need for more sustainable, healthier, and
livable cities has become a prominent concern. Planners and
decision-makers have proposed a shift to neighbourhood planning
principles centered on proximity to urban functions
\citep{pozoukidou15MinuteCityDecomposing2021}. In this context, the
15-Minute City is now in the public spotlight
\citep{logan_xminute_2022, moreno_introducing_2021}. The 15-Minute City
was introduced in \citet{moreno_introducing_2021} along with four
dimensions: density (in terms of people per km\(^2\)), diversity
(including mixed-land use and diversity of people), the temporal and
spatial proximity to essential services, and digitalisation
\citep[q.v.][]{cerveroTravelDemand3Ds1997}. However, the framework
presented has been criticized within academic and planning circles
\citep[e.g.,][]{guzmanProximityEnoughCritical2024, mouratidisTimeChallenge15minute2024}
among other things for shortfalls in terms of addressing pre-existing
structural forces and individual characteristics that drive inequalities
and influence who benefits or could benefit the most from such an
approach
\citep{dimarino15minuteCityConcept2023, willberg15minuteCityAll2023}. A
15-Minute City for all people is an aspirational goal, but does not
fully confront existing mobility and accessibility inequalities in need
of redress. Without directed and context-specific solutions, this
popular concept risks becoming an empty city branding exercise.
\citep{pozoukidou15MinuteCityDecomposing2021, gowerPlanningInnovationCity2022}.

Reflective of the flexible and aspirational presentation of the
15-Minute City, the concept has been adopted by cities using a diverse
range of definitions and tools along with indistinct universal
approaches. A trail blazer in this respect was Portland (US) in the
Portland Plan \citep{city_of_portland_20minute_2010} of April 2012.
Adopted before the `15-Minute City' concept of today, this plan aimed to
foster inclusive city development based on prosperity, education,
health, and equity over a 30 year horizon. Central to the plan was the
promotion of neighbourhood self-sufficiency and connectivity to city
centers and centres of employment. The progress report describes a
high-level focus on equitable service delivery to all residents with
equity concerning topics related to racial equity and people with
disabilities \citep{portland_government_portland_2017}, similarly taking
on an ``all populations'' approach. Subsequently, other cities adopted
proximity-based goals using similarly neutral approaches. For example,
the 15-Minute City plan announced during the re-election campaign of
Paris mayor Anne Hidalgo in 2020 emphasized six key social functions
that should be easily accessible from any location
\citep{ville_de_paris_paris_2022}. These locations included: housing,
work, health care, groceries, education and leisure. The 15-Minute City
concept inspired language in the agendas of other cities in the Western
world such as Ottawa, in the Canadian context, who also adopted a
15-Minutes approach in their recent Official Plan
\citep{ville_dottawa_quartier_2021}.
\citet{teixeiraClassifying15minuteCities2024}'s global review finds that
the 15-Minute City concept is in early phases of implementation around
the world and the diverse range of definitions, strategies, and
instruments present a significant knowledge and implementation gap. In
other words: the 15-Minute City is aspirational but how do we get there
and who will benefit?

@@

In examining how to enable improved quality of life through urban built
environment, accessibility measures have become an increasingly popular
tool. These measures are a way to quantify the ease of reaching
destinations from a given point in space and have been used to examine
urban areas through just and sustainable city planning agendas
\citep{vale2023accessibility, handyAccessibilityIdeaWhose2020}. The
15-Minute City is an amenity-provision neighbourhood planning concept
aimed at enabling the creation of urban environments that enhance life
quality, making it well-suited for analysis using accessibility
methodologies \citep{guzmanProximityEnoughCritical2024}. Recent works
have applied accessibility measures to investigate the 15-Minute City
across different geographic scopes. For instance, in Naples, Italy
\citep{gaglioneUrbanAccessibility15minute2022}, Barcelona, Spain
\citep{graellsCityCitiesMeasuring2021}, Vancouver, Canada
\citep{hosford15minuteCityReach2022}, and in urban areas across Europe
\citep{vale2023accessibility}. The ``cumulative opportunity'' measure
has been applied in many 15-Minute City examinations. This measure
quantifies how many destinations can be reached from a point in space
within a given travel time threshold , pairing well with normative
examinations \citep[see][]{paez_measuring_2012} that use travel time
thresholds like x-minute cities {[}\citet{logan_xminute_2022}.
Furthermore, the cumulative opportunity measure is widely appreciated
for its intuitive computation and popularity among practitioners
\citep{handyAccessibilityIdeaWhose2020, handyMeasuringAccessibilityExploration1997, chengInvestigatingWalkingAccessibility2019}.
However, accessibility measures other than the cumulative opportunity
have also been applied, reflecting the diversity of measures in the
literature \citep{guzmanProximityEnoughCritical2024}.

@@

Classification algorithms also show promise as a tool in the assessment
of the 15-Minute City, as they have been useful in identifying mobility
and spatial typologies within the transportation planning literature.
Often, these algorithms take the form of machine learning approaches: in
transportation, the use of Self-Organizing Maps (SOM) was pioneered by
Delmelle \citeyearpar{delmelle_spatial_2012} and has been used to group
U.S. neighbourhoods by minimizing dissimilarity in attributes
\citep{delmelleDifferentiatingPathwaysNeighborhood2017} and to classify
individuals' travel patterns according to the dissimilarity in mobility
attributes and uses the Decision Tree algorithm to partition the data
into interpretable classifications \citep{victorianoTimeSpaceMoney2020}.
Other approaches include the use of spatially constrained multivariate
clustering, to develop urban form typologies related to the 15-Minute
City \citep{burkeGeospatialAnalysisFramework2022} and the use of k-means
to analyse travel behaviour and classify metro stations based on
mobility patterns\citep{ganUnderstandingUrbanMobility2020}.

@@

Rather than focusing on \emph{all} destinations, it is valuable to
examine those related to caring activities. This framing offers a
feminist perspective on urban functions that matter from a care
perspective, and connects well with the 15-Minute City concept. Caring
activities, which fulfill the physical, psychological, and emotional
needs of others, are among the most essential and fundamental activities
in society \citep{ilo_care_2018}. Yet, they are the most unequal,
undervalued, and even devalued activities worldwide. Conventionally,
caring activities have been borne on women's shoulders
\citep{haydenGrandDomesticRevolution1982, hochschild_second_2012}.
According to \citet{ilo_care_2018}, women and girls perform more than
three-quarters of the total amount of unpaid care worldwide, a gender
gap that varies geographically \citep{RN5}. This unequal share of caring
responsibilities leads to multifaceted gendered differences: in career
development, profession selection, contract type, pay gap, and time
poverty, as recognized by various international organisations
\citep{EIGE2016, ilo_care_2018}. In terms of spatial and transportation
planning, almost one third of daily trips are for care purposes
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
From this research motivation, Sánchez de Madariaga coined the term
Mobility of Care in 2013 to refer to all travel required to sustain the
needs of a household, such as grocery shopping, escorting children,
travelling to health appointments, and running errands
\citep{sanchezdemadariagaMobilityCareIntroducing2013}. While an
undercurrent of research had examined these unique household-serving
trips over the decades, her seminal work was the first to consider all
these trips as one category and demonstrate how mobility of care is a
significant proportion of daily travel.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{unicode-math}
\usepackage{hyperref}
\def\tightlist{}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Towards completely caring 15-minute neighbourhoods},
  pdfauthor={Anastasia Soukhov; Léa Ravensbergen; Lucía Mejía Dorantes; Antonio Páez},
  pdfkeywords={Accessibility, Mobility of Care, Gender, 15-minute
neighbourhood, Care Complete Neighbourhoods},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Towards completely caring 15-minute neighbourhoods}
\author[1]{Anastasia Soukhov%
\corref{cor1}%
}
 \ead{soukhoa@mcmaster.ca} 
\author[1]{Léa Ravensbergen%
%
}
 \ead{ravensbl@mcmaster.ca} 
\author[2]{Lucía Mejía Dorantes%
%
}
 \ead{lucia.mejia.dorantes@gmail.com} 
\author[1]{Antonio Páez%
%
}
 \ead{paezha@mcmaster.ca}

@@

The ``15-Minute City'' concept has been embraced by global leaders to
promote human-scale neighbourhoods with transport and land-use designs
that support short trips to daily necessities. This paper bridges the
15-Minute City to ``Mobility of Care'', a framework that foregrounds
travel to care destinations, travel done predominately by women. This
focus contrasts the more commonly studied travel to employment and
leisure destinations. While the 15-Minute City concept is flexible
enough to consider all destination types, gendered examinations are
relatively lacking in the literature, and little research to date has
focused explicitly on care destinations. This gap is addressed in this
paper by identifying which areas in the city of Hamilton, Canada are
`caring 15-minute neighbourhoods'. To do so, a database of care
destinations was compiled to estimate the number (using the cumulative
opportunity accessibility measure) and diversity of mix (using the
entropy measure) of care destinations within a 15-minute walk from
residential parcels. Using data-driven machine learning techniques
(Self-Organizing Maps and Decision Trees), neighborhoods were classified
into `caring 15-minute neighbourhood' typologies that are examined
across residential socio-economic profiles. Our results suggest that the
majority of caring 15-minute neighbourhoods are in the urban core, where
more lower-income households reside. In contrast, not caring 15-minute
neighbourhoods in higher-income peripheral areas. Areas that make good
candidates for urban policy intervention are identified, and the
implications of enhancing 15-minute walkable caring access are discussed
in relation to gender mainstreaming in transportation planning.

Keywords: 15-Minute City; 15-Minute Neighbourhoods; Chrono-Urbanism;
Accessibility; Mobility of Care; Cities of Proximity; Inequality; Gender
Gap

@@

The term 15-Minute City is relatively new, but the concept is not. Its
newfound fame under the moniker of chrono-urbanism comes to complement
long-standing efforts to foster community and local travel to amenities
in neighbourhood planning practice. As recent examples, planned
neighbourhoods dominated post-WWII built urban form: the Neighbourhood
Unit Concept in the western world
\citep{brodyNeighbourhoodUnitConcept2013} and a parallel concept of the
``mikrorayons'' neighbourhoods in the Soviet Union
\citep{kissfazekasCircleParadigms15minute2022}. Along different
dimensions, these planning forms have been extensively critiqued
\citep{talenSocialSciencePlanned2017}. In this respect, neighbourhood
planning approaches such as the 15-Minute City concept offer an
opportunity for bottom-up approaches that leverage co-creation tools and
meaningful resident participation to achieve equitable and just
neighbourhood forms \citep{mahmoudCocreationPathwayUrban2021}. How to
prescribe equitable urban forms that `enable' contact with social
opportunities (instead of `engendering it' from the top-down) is still a
matter for debate. Questions remain, including: what destinations really
matter? What mode and travel time threshold is most appropriate? And who
benefits? To be certain, these dimensions are under discussion among
proximity-based planners who use accessibility-based tools
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022, guzmanProximityEnoughCritical2024}.
In this way, the 15-Minute City could be understood to be a normative
cumulative opportunity accessibility measure
\citep{paez_measuring_2012}.

From another perspective, feminist geographers, urban planners and other
researchers agree that urban structure influences travel, but city
planning falls short in planning for a neutral identity. In contrast,
cities ought to be planned according to the multiple identities of the
inhabitants
\citep{vacchelli_towards_2018, urbandevelopmentviennaGenderMainstreamingUrban2013}.
Researchers have demonstrated how mobility patterns differ according to
gender
\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990, neto_gender_2015}.
The causes for this are varied, and related to gendered social norms and
cultural factors that play a role in how, where, with whom, what for,
and how far we move. Women tend to travel shorter distances and more
frequently \citep{roorda_triap_2010, morency_distance_2011}, and conduct
more activities related to care no matter their stage of life
\citep{ilo_care_2018, garciaromanGenderDifferencesTime2022}. Combined,
these findings indicate that the spatial organization of activities has
a different impact according to gender, and therefore the 15-Minute City
concept should not be gender blind.

This work builds a theoretical bridge between the 15-Minute City and
Mobility of Care concepts to answer questions about what destinations
matter and who benefits from proximity planning. The Mobility of Care
term was coined in
\citet{sanchezdemadariagaMobilityCareIntroducing2013}; it refers to all
travel required to sustain the needs of a household, such as grocery
shopping, escorting children, travelling to health appointments, and
running errands. While decades of research have examined types of
household-serving trips (e.g., shopping trips, escorting to school),
\citet{sanchezdemadariagaMobilityCareIntroducing2013} was the first to
consider them all as one category. In doing so, care-related mobility
can be shown to comprise a significant proportion of daily travel,
characterizing approximately 30\% of adults' daily trips
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
We argue that Mobility of Care is especially relevant to the 15-Minute
City, as care trips are necessary for \emph{all} people and they most
often occur at the local level. Moreover, care trips are of particular
gendered and socio-economic importance, as women, and especially those
within lower income households, complete the majority of care trips
\citep{ravensbergen2023exploratory}. Despite care travel's importance in
advancing gender equality -and equity more broadly-, Mobility of Care
and access to care destinations have been under-examined relative to
employment destinations. This is especially pertinent in accessibility
research that has largely focused on employment points of interest e.g.,
\citep{farberOntarioLineSocioeconomic2019, duarteInfluenceJobAccessibility2023, ryanAccessibilitySpaceTime2023}.

@@

In the last decades, the need for more sustainable, healthier, and
livable cities has become a prominent concern. Planners and
decision-makers have proposed a shift to neighbourhood planning
principles centered on proximity to urban functions
\citep{pozoukidou15MinuteCityDecomposing2021}. In this context, the
15-Minute City is now in the public spotlight
\citep{logan_xminute_2022, moreno_introducing_2021}. The 15-Minute City
was introduced in \citet{moreno_introducing_2021} along with four
dimensions: density (in terms of people per km\(^2\)), diversity
(including mixed-land use and diversity of people), the temporal and
spatial proximity to essential services, and digitalisation
\citep[related concepts in][]{cerveroTravelDemand3Ds1997}. However, the
framework presented has been criticized within academic and planning
circles
\citep[e.g.,][]{guzmanProximityEnoughCritical2024, mouratidisTimeChallenge15minute2024}
among other things for shortfalls in terms of addressing pre-existing
structural forces and individual characteristics that drive inequalities
and influence who benefits or could benefit the most from such an
approach
\citep{dimarino15minuteCityConcept2023, willberg15minuteCityAll2023}. A
15-Minute City for all people is an aspirational goal, but does not
fully confront existing mobility and accessibility inequalities in need
of redress. Without directed and context-specific solutions, this
popular concept risks becoming an empty city branding exercise.
\citep{pozoukidou15MinuteCityDecomposing2021, gowerPlanningInnovationCity2022}.

Reflective of the flexible and aspirational presentation of the
15-Minute City, the concept has been adopted by cities using a diverse
range of definitions and tools along with indistinct universal
approaches. A trail blazer in this respect was Portland (U.S.) in the
Portland Plan \citep{city_of_portland_20minute_2010} of April 2012.
Adopted before the `15-Minute City' concept of today, this plan aimed to
foster inclusive city development based on prosperity, education,
health, and equity over a 30 year horizon. Central to the plan was the
promotion of neighbourhood self-sufficiency and connectivity to city
centers and centres of employment. The progress report describes a
high-level focus on equitable service delivery to all residents with
equity concerning topics related to racial equity and people with
disabilities \citep{portland_government_portland_2017}, similarly taking
on an ``all populations'' approach. Subsequently, other cities adopted
proximity-based goals using similarly neutral approaches. For example,
the 15-Minute City plan announced during the re-election campaign of
Paris mayor Anne Hidalgo in 2020 emphasized six key social functions
that should be easily accessible from any location
\citep{ville_de_paris_paris_2022}. These locations included: housing,
work, health care, groceries, education and leisure. The 15-Minute City
concept inspired language in the agendas of other cities in the Western
world such as Ottawa, in the Canadian context, who also adopted a
15-Minutes approach in their recent Official Plan
\citep{ville_dottawa_quartier_2021}.
\citet{teixeiraClassifying15minuteCities2024}'s global review finds that
the 15-Minute City concept is in early phases of implementation around
the world and the diverse range of definitions, strategies, and
instruments present a significant knowledge and implementation gap. In
other words: the 15-Minute City is aspirational but how do we get there
and who will benefit?

@@

In examining how to enable improved quality of life through urban built
environment, accessibility measures have become an increasingly popular
tool. These measures are a way to quantify the ease of reaching
destinations from a given point in space and have been used to examine
urban areas through just and sustainable city planning agendas
\citep{vale2023accessibility, handyAccessibilityIdeaWhose2020}. The
15-Minute City is an amenity-provision neighbourhood planning concept
aimed at enabling the creation of urban environments that enhance life
quality, making it well-suited for analysis using accessibility
methodologies \citep{guzmanProximityEnoughCritical2024}. Recent works
have applied accessibility measures to investigate the 15-Minute City
across different geographic scopes. For instance, in Naples, Italy
\citep{gaglioneUrbanAccessibility15minute2022}, Barcelona, Spain
\citep{graellsCityCitiesMeasuring2021}, Vancouver, Canada
\citep{hosford15minuteCityReach2022}, and in urban areas across Europe
\citep{vale2023accessibility}. The ``cumulative opportunity'' measure
has been applied in many 15-Minute City examinations. This measure
quantifies how many destinations can be reached from a point in space
within a given travel time threshold, pairing well with normative
examinations \citep[see][]{paez_measuring_2012} that use travel time
thresholds like x-minute cities \citep{logan_xminute_2022}. Furthermore,
the cumulative opportunity measure is widely appreciated for its
intuitive computation and popularity among practitioners
\citep{handyAccessibilityIdeaWhose2020, handyMeasuringAccessibilityExploration1997, chengInvestigatingWalkingAccessibility2019}.
However, accessibility measures other than the cumulative opportunity
have also been applied, reflecting the diversity of measures in the
literature \citep{guzmanProximityEnoughCritical2024}.

@@

Classification algorithms also show promise as a tool in the assessment
of the 15-Minute City, as they have been useful in identifying mobility
and spatial typologies within the transportation planning literature.
Often, these algorithms take the form of machine learning approaches: in
transportation, the use of Self-Organizing Maps (SOM) was pioneered by
Delmelle \citeyearpar{delmelle_spatial_2012}. It has been used to group
U.S. neighbourhoods by minimizing dissimilarity in attributes
\citep{delmelleDifferentiatingPathwaysNeighborhood2017}. SOM has also
been used to classify individuals' travel patterns according to the
dissimilarity in mobility attributes along with the Decision Tree
algorithm to partition the data into interpretable classifications
\citep{victorianoTimeSpaceMoney2020}. Other approaches include the use
of spatially constrained multivariate clustering, to develop urban form
typologies related to the 15-Minute City
\citep{burkeGeospatialAnalysisFramework2022} and the use of k-means to
analyse travel behaviour and classify metro stations based on mobility
patterns \citep{ganUnderstandingUrbanMobility2020}.

@@

Rather than focusing on \emph{all} destinations, it is valuable to
examine those related to caring activities. This framing offers a
feminist perspective on urban functions that matter from a care
perspective, and connects well with the 15-Minute City concept. Caring
activities, which fulfill the physical, psychological, and emotional
needs of others, are among the most essential and fundamental activities
in society \citep{ilo_care_2018}. Yet, they are the most unequal,
undervalued, and even devalued activities worldwide. Conventionally,
caring activities have been borne on women's shoulders
\citep{haydenGrandDomesticRevolution1982, hochschild_second_2012}.
According to \citet{ilo_care_2018}, women and girls perform more than
three-quarters of the total amount of unpaid care worldwide, a gender
gap that varies geographically \citep{RN5}. This unequal share of caring
responsibilities leads to multifaceted gendered differences: in career
development, profession selection, contract type, pay gap, and time
poverty, as recognized by various international organisations
\citep{EIGE2016, ilo_care_2018}. In terms of spatial and transportation
planning, almost one third of daily trips are for care purposes
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
From this research motivation, Sánchez de Madariaga coined the term
Mobility of Care in 2013 to refer to all travel required to sustain the
needs of a household, such as grocery shopping, escorting children,
travelling to health appointments, and running errands
\citep{sanchezdemadariagaMobilityCareIntroducing2013}. While an
undercurrent of research had examined these unique household-serving
trips over the decades, her seminal work was the first to consider all
these trips as one category and demonstrate how Mobility of Care is a
significant proportion of daily travel.
